---
title: "covid19_project"
author: "Peter Mehler"
date: "3/18/2020"
output: html_document
---
data set: https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series

```{r}
confirmed_cases <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv")

deaths <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv")

recovered <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv")
```

#TO DO:
Slope analysis / Comparison?
Make multiple province countries functional with clean_data()
Plot Predictive values extending from log of reported infections
Look at Active cases are calculated as, Active cases = Confirmed cases - Deaths reported - Recovered cases
Clean up to make it more report oriented

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

```{r}
us <- confirmed_cases %>% 
  filter(Province.State=="New York" | Province.State=="Washington" |Province.State=="California" | Province.State=="New Jersey" | Province.State=="Massachusetts" | Province.State=="Florida") %>% 
  gather(key = X1.22.20, value = cases, 5:names(confirmed_cases)[length(names(confirmed_cases))]) %>% 
  mutate(date = sub("X", "", X1.22.20)) %>% 
  select(-X1.22.20) %>% 
  mutate(date = as.Date(date,format="%m.%d.%y"))
```

```{r}
us_plot <- ggplot(data = us, aes(x=date, y=cases, group = Province.State, color = Province.State)) +
  labs(title = "Number of Cases in Top 6 Most Infected US States") +
  geom_line() +
  xlim(as.Date("3.8.20",format="%m.%d.%y"),NA)

us_plot
```
# Clean Hubei Data
```{r}
# region == 1 means country, region == 2 means province
clean_data <- function(name, region){
   if (region==1){
    temp_deaths <- deaths %>% 
      filter(Country.Region==name) %>% 
      #Convert to long format
      gather(key = X1.22.20, value = deaths, 5:names(confirmed_cases)[length(names(confirmed_cases))]) %>%
      mutate(date = sub("X", "", X1.22.20)) %>%  #get rid of the X prefix
      select(-X1.22.20) %>% #delete old column
      mutate(date = as.Date(date,format="%m.%d.%y")) %>%  #turn it into a date object
      select(deaths, date)
    
    temp_recovered <- recovered %>% 
      filter(Country.Region==name) %>% 
      gather(key = X1.22.20, value = recovered, 5:names(confirmed_cases)[length(names(confirmed_cases))]) %>% 
      mutate(date = sub("X", "", X1.22.20)) %>% 
      select(-X1.22.20) %>% 
      mutate(date = as.Date(date,format="%m.%d.%y")) %>% 
      select(recovered, date)
      
    temp_cases <- confirmed_cases %>% 
      filter(Country.Region==name) %>% 
      gather(key = X1.22.20, value = cases, 5:names(confirmed_cases)[length(names(confirmed_cases))]) %>% 
      mutate(date = sub("X", "", X1.22.20)) %>% 
      select(-X1.22.20) %>% 
      mutate(date = as.Date(date,format="%m.%d.%y"))
    
    temp_cases <- left_join(temp_cases, temp_deaths, "date")
    temp_cases <- left_join(temp_cases, temp_recovered, "date")
    
    temp_cases <- temp_cases %>%
      gather(key = type, value = amount, cases, recovered, deaths)
  
    return(temp_cases)
  }
  
  else{
    temp_deaths <- deaths %>% 
      filter(Province.State==name) %>% 
      gather(key = X1.22.20, value = deaths, 5:names(confirmed_cases)[length(names(confirmed_cases))]) %>% 
      mutate(date = sub("X", "", X1.22.20)) %>% 
      select(-X1.22.20) %>% 
      mutate(date = as.Date(date,format="%m.%d.%y")) %>% 
      select(deaths, date)
    
    temp_recovered <- recovered %>% 
      filter(Province.State==name) %>% 
      gather(key = X1.22.20, value = recovered, 5:names(confirmed_cases)[length(names(confirmed_cases))]) %>% 
      mutate(date = sub("X", "", X1.22.20)) %>% 
      select(-X1.22.20) %>% 
      mutate(date = as.Date(date,format="%m.%d.%y")) %>% 
      select(recovered, date)
      
    temp_cases <- confirmed_cases %>% 
      filter(Province.State==name) %>% 
      gather(key = X1.22.20, value = cases, 5:names(confirmed_cases)[length(names(confirmed_cases))]) %>% 
      mutate(date = sub("X", "", X1.22.20)) %>% 
      select(-X1.22.20) %>% 
      mutate(date = as.Date(date,format="%m.%d.%y"))
    
    temp_cases <- left_join(temp_cases, temp_deaths, "date")
    temp_cases <- left_join(temp_cases, temp_recovered, "date")
    
    temp_cases <- temp_cases %>%
      gather(key = type, value = amount, cases, recovered, deaths)
  
    return(temp_cases)
  }
}
```
# CHOOSE YOUR COUNTRY OR PROVINCE
Only works for Countries without several provinces and single provinces \n
region == 1 means country, region == 2 means province
```{r}
hubei <- clean_data("Hubei", 2)
```
```{r}
italy <- clean_data("Italy", 1)
```
```{r}
s.korea <- clean_data("Korea, South", 1)
```
```{r}
ny <- clean_data("New York", 2)
```
```{r}
iran <- clean_data("Iran", 1)
```
```{r}
spain <- clean_data("Spain",1)
```


```{r}
# Works for countries with several provinces
us_total <- confirmed_cases %>% 
  filter(Country.Region=="US") %>% 
  gather(key = X1.22.20, value = cases, 5:names(confirmed_cases)[length(names(confirmed_cases))]) %>% 
  mutate(date = sub("X", "", X1.22.20)) %>% 
  select(-X1.22.20) %>% 
  mutate(date = as.Date(date,format="%m.%d.%y")) %>%
  gather(key = type, value = amount, cases) %>% 
  group_by(date) %>% 
  summarize(amount = sum(amount),
            Country.Region = "US")
```

```{r, include=FALSE}
hubei %>% 
  filter(type=="cases") %>% 
  ggplot(aes(x=date, y=amount, group = 1)) +
  labs(title = "hubei province, red line indicates Wuhan shutdown") +
  geom_line() + 
  geom_vline(xintercept = as.Date("1.23.20",format="%m.%d.%y"), 
                color = "red", size=1) +
  xlim(as.Date("1.22.20",format="%m.%d.%y"),NA)
  
#Wuhan Lockdown January 23rd
```
```{r}
ggplot(data = hubei, aes(date, amount, fill = type)) +
  geom_bar(position="stack", stat = "identity") +
  labs(title = "Cases, Deaths, and Recoveries in Hubei Province Through Time",
       x = "Date",
       y = "Amount") +
  theme(legend.title=element_blank())
```

# Italy Plot

```{r, include=FALSE}
italy %>% 
  filter(type=="cases") %>% 
  ggplot(aes(x=date, y=amount, group = 1)) +
  labs(title = "italy", caption = "caption" ) +
  geom_line() + 
  geom_vline(xintercept = as.Date("3.8.20",format="%m.%d.%y"), 
                color = "red", size=1) +
  xlim(as.Date("3.8.20",format="%m.%d.%y"),NA) +
  #ylim(0,32000) +
  geom_smooth() +
  #geom_smooth(method = lm) +
  theme_minimal()

# Perhaps there's a better way to calculate slope, Trying to avoid manual entry
# slope = round(lm(temp$date~temp$amount)$coefficients[2],4)
# 
# as.double(slope)*10000
#Italy Lockdown March 8th
```
```{r}
ggplot(data = italy, aes(date, amount, fill = type)) +
  geom_bar(position="stack", stat = "identity") +
  theme(legend.title=element_blank()) +
  labs(title = "Cases, Deaths, and Recoveries in Italy Through Time",
       x = "Date",
       y = "Amount")
```
```{r}
require(gridExtra)
plot1 <- ggplot(data = italy, aes(date, amount, fill = type)) +
  geom_bar(position="fill", stat = "identity") +
  labs(title = "Percentages of Case Statistics Italy", y="Percent", x="Date") +
  theme(legend.title=element_blank())
plot2 <- ggplot(data = hubei, aes(date, amount, fill = type)) +
  geom_bar(position="fill", stat = "identity") +
  labs(title = "Percentages of Case Statistics Hubei", y="Percent", x="Date") +
  theme(legend.title=element_blank())

grid.arrange(plot1, plot2, ncol=2)
```


```{r}
# Row combine all datasets from Italy, SK, Hubei, New York
combined.countries <- rbind(hubei, italy, s.korea, ny, iran, spain)

combined.countries <- combined.countries %>%
  filter(type=="cases") %>%   # Have filtered by cases
  filter(amount>400) %>%    # Filter cases > 400
  group_by(Country.Region) %>% 
  mutate(date = as.integer(format(date, "%j"))) %>% 
  mutate(date = date - date[1]) # Change date to days since cases > 400
```
```{r}
ggplot(data = combined.countries, aes(date, amount, color = Country.Region)) +
  #geom_line() +
  # geom_vline(xintercept = 1, color = "red", size=.25) + # Wuhan locks down
  # geom_vline(xintercept = 11, color = "green", size=.25) + # Italy locks down
  # geom_vline(xintercept = -1, color = "blue", size=.25) + # South Korea Socially Isolates
  # geom_vline(xintercept = 8, color = "purple", size=.25) + # NY State orders lockdown
  labs(x="Days since number of cases reached 400", 
       y="Cumulative Cases", 
       title = "Comparison of Cumulative Cases by Country",
       subtitle = "Beginning When Number of Infections is Above 400",
       caption = "US represents only NY data and China represents only Hubei data") +
  xlim(0,NA) +
  ylim(0,NA) +
  geom_smooth(size=.75) +
  theme(legend.title=element_blank())
                        

#Majority of people of South Korea decide to socially isolate 2.20.20 
#Italy Lockdown 3.8.20
#Wuhan Lockdown 1.23.20
#NY State orders lockdown
```
```{r}
ggplot(data = combined.countries, aes(date, amount, color = Country.Region)) +
  geom_line() +
  labs(x="Days since number of cases reached 400", 
       y="Cumulative Cases", 
       title = "Exponential Projections of Cumulative Cases by Country",
       subtitle = "Beginning When Number of Infections is Above 400",
       caption = "US represents only NY data and China represents only Hubei data") +
  xlim(NA,50) +
  ylim(0,NA) +
  #geom_smooth(size=.75) +
  theme(legend.title=element_blank()) +
  scale_y_log10()
  #stat_smooth(method="lm",fullrange=TRUE) 
```

```{r}
# Row combine all datasets from Italy, SK, Hubei, New York
combined.countries <- rbind(hubei, italy, s.korea, ny, iran)

combined.countries <- combined.countries %>%
  filter(type=="deaths") %>%   # Have filtered by deaths
  filter(amount>0) %>%    # Filter cases > 0
  group_by(Country.Region) %>% 
  mutate(date = as.integer(format(date, "%j"))) %>% 
  mutate(date = date - date[1]) # Change date to days since first death
```
```{r}
ggplot(data = combined.countries, aes(date, amount, color = Country.Region)) +
  #geom_line() +
  labs(x="Days since first Death", 
       y="Cumulative Deaths", 
       title = "Comparison of Cumulative Deaths by Country Beginning After First Death", 
       caption = "US represents only NY data and China represents only Hubei data") +
  ylim(0,NA) +
  geom_smooth() +
  theme(legend.title=element_blank())

```
```{r}
# Row combine all datasets from Italy, SK, Hubei, New York
combined.countries <- rbind(hubei, italy, s.korea, iran)

combined.countries <- combined.countries %>%
  filter(type=="deaths") %>%   # Have filtered by deaths
  filter(amount>0) %>%    # Filter cases > 0
  group_by(Country.Region) %>% 
  mutate(date = as.integer(format(date, "%j"))) %>% 
  mutate(date = date - date[1]) # Change date to days since first death
```
```{r}
ggplot(data = combined.countries, aes(date, amount, color = Country.Region)) +
  #geom_line() +
  labs(x="Days since first Death", 
       y="Cumulative Deaths", 
       title = "Exponential Projections of Cumulative Deaths by Country Beginning After First Death", 
       caption = "US represents only NY data and China represents only Hubei data") +
  ylim(0,NA) +
  #geom_smooth() +
  theme(legend.title=element_blank()) +
  scale_y_log10() +
  stat_smooth(method="lm",fullrange=TRUE)
```


Finding regression line for different subsets of data:

Source: https://stackoverflow.com/questions/51355303/extract-slope-of-multiple-trend-lines-from-geom-smooth